Capstone Movielens Report

### 2025-12-10 reports/capstone-movielens.site/05.appendix.Rmd
Function described: `loess_de`.

@@

Builds the *User+Movie+Genre+Year Effect (UMGYDE) Model* based on the *Day General Effect* data structure returned by the [calc_date_general_effect](#func.calc_date_general_effect) or [calc_date_general_effect.cv](#func.calc_date_general_effect.cv) function (both described above) and passed in the [date_general_effect](#func.calc_UMGY_SmoothedDay_effect.args.date_general_effect) argument.

@@

@@

@@

@@

@@

::: {.noteblock data-latex=""}
The complete source code of the functions described in this section is available in the  [Regularization](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L224) section of the [UMGYD-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L3) script on _GitHub_.
:::

@@

Under the hood, it calls internally the [calc_date_general_effect.cv](#func.calc_date_general_effect.cv) and [calc_UMGY_effect](#func.calc_UMGY_effect) functions (both described above in this section).

>>>>

#### [loess_de](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L69) Function {#func.loess_de}
\

Using local fitting, fits a polynomial surface determined by the *Day General Effect* (hereafter *DG Effect* or *DGE* for short) information passed in the [de_bias.dat](#func.loess_de.args.de_bias.dat) argument.

::: {.noteblock data-latex=""}
In *this Project*, the *DGE* data object to be passed in the [de_bias.dat](#func.loess_de.args.de_bias.dat) argument is returned by the [calc_day_general_effect](#func.calc_day_general_effect) and [calc_day_general_effect.cv](#func.calc_day_general_effect.cv) functions (both described above).
:::

::: { #func.loess_de.usage .sidebar }
##### Usage
\

```{r eval=FALSE}
  fit <- day_gen_effect |> loess_de(degree, span)
```

::: {.noteblock data-latex=""}
In the code snippet above, the `loess_de` is called internally from the [calc_UMGY_SmoothedDay_effect](#func.calc_UMGY_SmoothedDay_effect) function described below.
:::
:::

::: { #func.loess_de.args .sidebar }
##### Arguments
\

::: { #func.loess_de.args.de_bias.dat .sidebar }
###### ***de_bias.dat***
\

Dataframe object representing the *Day General Effect* data.

:::
::: { #func.loess_de.args.degree .sidebar }
###### ***degree***
\

(Integer, $NA$ by default) The degree of the polynomials to be used (passed in the same-named argument to the internally called [stats::loess](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/loess) function).

::: {.noteblock data-latex=""}
When the value is $NA$ (default), $1$ is passed in the same-named argument to the internally called [stats::loess](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/loess) function. 
:::
:::
::: { #func.loess_de.args.span .sidebar }
###### ***span***
\

(Numeric, $NA$ by default ) The parameter $\alpha$, which controls the degree of smoothing (passed in the same-named argument to the internally called [stats::loess](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/loess) function).

::: {.noteblock data-latex=""}
When the value is $NA$ (default), $0.75$ is passed in the same-named argument to the internally called [stats::loess](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/loess) function. 
:::
:::
:::

::: { #func.loess_de.details .sidebar }
##### Details
\

This function is a wrapper for the [stats::loess](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/loess) `R` function, which is called internally to make a required fit.

::: {.noteblock data-latex=""}
This is a helper function intended for use in other helper functions, such as the [calc_UMGY_SmoothedDay_effect](#func.calc_UMGY_SmoothedDay_effect) described below in this section.
:::
:::

::: { #func.loess_de.value .sidebar }
##### Value
\

Object of class `loess` returned by the internally called function [stats::loess](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/loess).
:::

::: { #func.loess_de.code .sidebar }
##### Source Code
\

The source code of the [loess_de](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R#L71) function is shown below:
```{r eval=FALSE}
loess_de <- function(de_bias.dat, degree = NA, span = NA){
  if(is.na(degree)) degree = 1
  if(is.na(span)) span = 0.75
  loess(de ~ days, span = span, degree = degree, data = de_bias.dat)
}
```

::: {.noteblock data-latex=""}
The source code of the [loess_de](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L71) function is also available in the [Utility Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L3) section of the [UMGYD-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L71) script on _GitHub_.
:::
:::

@@

Builds the *User+Movie+Genre+Year Effect (UMGYDE) Model* based on the *Day General Effect* data structure returned by the [calc_day_general_effect](#func.calc_day_general_effect) or [calc_day_general_effect.cv](#func.calc_day_general_effect.cv) function (both described above) and passed in the [date_general_effect](#func.calc_UMGY_SmoothedDay_effect.args.date_general_effect) argument.

@@

Under the hood, this function internally calls the [calc_day_general_effect](#func.calc_day_general_effect) and [calc_UMGY_effect](#func.calc_UMGY_effect) functions (both described above in this section).

@@

Under the hood, this function internally calls the [calc_day_general_effect.cv](#func.calc_day_general_effect.cv) and [calc_UMGY_effect](#func.calc_UMGY_effect) functions (both described above in this section).

@@

  calc_day_general_effect.cv(lambda) |> calc_UMGY_effect()  
}
```

@@

### UMGYDE Model: Tuning `loess` Params
\

::: {.noteblock data-latex=""}
The complete source code of the functions described in this section is available in the  [Tuning `loess` Parameters](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L172) section of the [UMGYD-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L172) script on _GitHub_.
:::

@@

::: {.noteblock data-latex=""}
The complete source code of the functions described in this section is available in the  [Regularization](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L224) section of the [UMGYD-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L224) script on _GitHub_.
:::

@@

Under the hood, it calls internally the [calc_day_general_effect.cv](#func.calc_day_general_effect.cv) and [calc_UMGY_effect](#func.calc_UMGY_effect) functions (both described above in this section).
